\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-29 coqimpl.tex
First draft of challenges (replacing design space section)

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user
interactively inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking and overriding as in \ref{langdesign-req0}---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

@@

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": uninstantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.

@@

\textbf{Reasoning Needs Non-Overridable.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

@@

\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
We support ``extensible'' inductive types from
\ref{langdesign-req2} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it. It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}

@@

To construct a recursor as in \ref{langdesign-req3}, we need to create for each
constructor a handler (function) and aggregate them into a family, and thus that family (e.g. "neg_handler" in \cref{fig:plugin-example2}) can be considered as a large matching-case statement. We use "FRecursor" to declare a recursor (e.g. "neg") by specifying the motive, the aiming inductive type and the handler family. Once recursor is created it will look like any other function fields (e.g. "neg"). Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field. The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g. "uu"), then the inherited family (e.g. "neg_handler") will need to extend correspondingly (e.g. "neg_handler.uu"). If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".

@@

However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{langdesign-req3}), thus the information doesn't totally disappear but store elsewhere.  Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12". This concrete recursor will be used by our "FRecursor". For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g. "neg_handler") during compilation---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq. Note that, at this point of the example, "FRecursor" will only do type check. The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module: 
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module. Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.

@@

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{langdesign-req2}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff} because from the perspective of Coq they are just two terms declared in the module type.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only? My confusion after reading these paras is that I still don't know why partial recursors are needed.}\EDJreply{The issue is exactly, at that point of programming, for extensibility, there is no access to the concrete recursor. Concrete recursors also cannot be postulated as part of the interface because it breaks extensibility.}
In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.

@@

% However, to verify and prove the recursor and partial recursor and their
% computational rules (as in \ref{langdesign-req4}), our plugin has to
% compile the extensible family into a non-extensible Coq Module and define
% the recursors by instantiating the content of Coq module with concrete
% inductive types. We provide "Closing Fact" as a command to carry out this global reasoning,

@@

% This "Closing Fact" also can ease us from the necessity to let plugin
% generate proof for recursors and partial recursors and their
% computational rules (as in \ref{langdesign-req4}), and thus making
% plugin development easier.
%\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}\EDJreply{You are right. It does sound like a bad excuse. I rewrite the whole section now. Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma. \\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do). This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
% For example, the computation rule for a
% concrete recursor like "B.neg" and partial recursors in
% \cref{fig:plugin-example2} can be directly asserted by using "Closing
% Fact" and their proofs verified later during compilation.
%\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know. There must be some other ways in the literature. }\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}\EDJreply{Please check.}
% This ``Global Reasoning'' will be again useful in later examples.

@@

Thus to refer to them, we have to scope the defining tactic with
"MetaData" and our plugin will set up the appropriate environment. After
a "MetaData" block is closed, the plugin will aggregate the content into
the family being defined so that they can also be used when defining
the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure. This auto-insertion will practically look like hijack the original Coq command. I am not sure this kind of hijacking can happen.}
This can fulfill \ref{langdesign-req5}.

% To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of

@@

%   For example, currently it is not possible to define a
% vanilla inductive type dependent on our ``extensible'' inductive type
% due to the invisibility, unless we compile the whole family. Similarly,
% customized tactic expressions as required by \ref{langdesign-req5} cannot
% use lemmas or theorems in the fields of the current defining family as
% well.

% To remedy this, we support a "MetaData" command. This command
% bundles any original Coq primitive (in the current family
% context) into the compiled module type, and makes them visible to the
% following defining field. Doing so, we can define customized tactic as
% in \ref{langdesign-req5} that uses lemmas proved in the family.

>>>>

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user
interactively inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just an empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4" is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking and overriding as in \ref{chg:software-engineering}---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

@@

Note that, in our example, we include two
functors without arguments---this is due to a peculiarity of Coq's
"Include": uninstantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.
For example, "b_7" requires a module of type "a_6". Inside module "A",
when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat}
inside the current surrounding "A". Then when we
\mintinline{Coq}{Include b_7} without specifying the module parameter,
Coq will try to find fields in the surrounding module (i.e., module
"A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by
the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and 
thus maintains all the computational mechanism from Coq, resolving \ref{chg:software-engineering}.

@@

\textbf{Reasoning Needs Non-Overridable.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{chg:definition-relevant-reasoning}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

@@

\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
We support ``extensible'' inductive types from
\ref{chg:extensible-inductive-type} in Coq to some extent.\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort}\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it. It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}

@@

To construct a recursor as in \ref{chg:extensible-inductive-type}, we need to create for each
constructor a handler (function) and aggregate them into a family, and thus that family (e.g. "neg_handler" in \cref{fig:plugin-example2}) can be considered as a large matching-case statement. We use "FRecursor" to declare a recursor (e.g. "neg") by specifying the motive, the aiming inductive type and the handler family. Once recursor is created it will look like any other function fields (e.g. "neg"). Here in \cref{fig:plugin-example2} we apply "neg" to "tt" as an exemplar field. The inheritance of recursor is mainly delegated by the inheritance of the handler family---if the user extend with new constructor (e.g. "uu"), then the inherited family (e.g. "neg_handler") will need to extend correspondingly (e.g. "neg_handler.uu"). If not, the plugin will error with ``Non-exhaustiveness Pattern'' when inheriting the field "neg".

@@

However, we later still need the concrete eliminator "b" for doing type-checking (as in \ref{chg:extensible-inductive-type}), thus the information doesn't totally disappear but store elsewhere.  Here we show one exemplar extraction of the recursor to \mintinline{Coq}{Set} as module "b_rec_12". This concrete recursor will be used by our "FRecursor". For example, in \cref{fig:plugin-example2}, "FRecursor" command will feed Coq a module "v_33_34" for doing type-checking with specified handlers (e.g. "neg_handler") during compilation---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq. Note that, at this point of the example, "FRecursor" will only do type check. The compilation of "FRecursor" into a component of the compiled module only happens when closing the whole family, by inserting the following into the compiled module: 
\begin{minted}{Coq}
  Definition neg := b_rec (fun _ => b) neg_handler.tt neg_handler.ff.
\end{minted}
where "b_rec" and "b" are the corresponding recursor and the (vanilla) inductive type in that compiled module. Thus a "FRecursor" field is unlike other fields, even though it is exposed as a mundane function for the following fields.

@@

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{chg:extensible-inductive-type}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff} because from the perspective of Coq they are just two terms declared in the module type.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only? My confusion after reading these paras is that I still don't know why partial recursors are needed.}\EDJreply{The issue is exactly, at that point of programming, for extensibility, there is no access to the concrete recursor. Concrete recursors also cannot be postulated as part of the interface because it breaks extensibility.}
In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.

@@

% However, to verify and prove the recursor and partial recursor and their
% computational rules (as in \ref{chg:definition-relevant-reasoning}), our plugin has to
% compile the extensible family into a non-extensible Coq Module and define
% the recursors by instantiating the content of Coq module with concrete
% inductive types. We provide "Closing Fact" as a command to carry out this global reasoning,

@@

% This "Closing Fact" also can ease us from the necessity to let plugin
% generate proof for recursors and partial recursors and their
% computational rules (as in \ref{chg:definition-relevant-reasoning}), and thus making
% plugin development easier.
%\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}\EDJreply{You are right. It does sound like a bad excuse. I rewrite the whole section now. Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma. \\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do). This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
% For example, the computation rule for a
% concrete recursor like "B.neg" and partial recursors in
% \cref{fig:plugin-example2} can be directly asserted by using "Closing
% Fact" and their proofs verified later during compilation.
%\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know. There must be some other ways in the literature. }\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}\EDJreply{Please check.}
% This ``Global Reasoning'' will be again useful in later examples.

@@

Thus to refer to them, we have to scope the defining tactic with
"MetaData" and our plugin will set up the appropriate environment. After
a "MetaData" block is closed, the plugin will aggregate the content into
the family being defined so that they can also be used when defining
the following fields of the family.\YZ{It reads like this MetaData wrapping could be automatically inserted, no?}\EDJreply{Not sure. This auto-insertion will practically look like hijack the original Coq command. I am not sure this kind of hijacking can happen.}
This can fulfill \ref{chg:software-engineering}.

% To fulfill \ref{chg:software-engineering}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of

@@

%   For example, currently it is not possible to define a
% vanilla inductive type dependent on our ``extensible'' inductive type
% due to the invisibility, unless we compile the whole family. Similarly,
% customized tactic expressions as required by \ref{chg:software-engineering} cannot
% use lemmas or theorems in the fields of the current defining family as
% well.

% To remedy this, we support a "MetaData" command. This command
% bundles any original Coq primitive (in the current family
% context) into the compiled module type, and makes them visible to the
% following defining field. Doing so, we can define customized tactic as
% in \ref{chg:software-engineering} that uses lemmas proved in the family.

### 2022-06-29 lang-design.tex
First draft of challenges (replacing design space section)

\subsection{Design Space}\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).  I think most readers will benefit from a recap of this encoding.\\
Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
}

When adapting Family Polymorphism into dependent types, we choose to
focus only on the essence of family (and inheritance)
structure in \citet{zm2017}, and thus a lot of unrelated features, like
Interface, will be removed. In this case, it will look like module with
late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
need it to have good compatibility with inductive types, because we
don't want to retain the idiomatic Coq programming experience
%\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness. I rephrase it now, please check.}
(and its tactic programming).

Let's start with a basic example---STLC and its extension in
\cref{fig:STLC-example}---to consider what kind of features are required
and how much of them can be supported by family polymorphism.
% add the example code here

\begin{figure}[!htb]
  \begin{minipage}[t]{0.47\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=||]{Coq}
Family STLC.|\YZ{Does minted support adding extra keywords such as 'Family' and 'extends'?}|
Inductive ty : Set :=
  | unit : ty
  | arrow : ty -> ty -> ty.
Inductive term : Set := 
  | var : id -> term 
  | lam : id -> term -> term ...
Fixpoint subst 
  : id -> term -> term := ...
Inductive has_type 
  : context -> term -> type := ...
Proposition subst_lemma :
  has_type (Γ; x ↦ A) t T ->
  has_type Γ u A ->
  has_type Γ (subst x u t) T.
Inductive step : term -> term -> Prop 
  := ...
(* ... and more, end with type safety *)
EndFamily.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.47\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family STLC_bool extends STLC.
Extend Inductive ty : Set :=
  | bool : ty.

Extend Inductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term
Extend Fixpoint subst (* Need to handle new term *)
  : id -> term -> term := ...
Extend Inductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
Extend Proposition subst_lemma :
  ... (* Need to prove extra cases *)

@@

First and foremost, we don't want to throw away some basic good
properties of Coq. \textlabel{Req~(0A)}{langdesign-req0}~\textbf{we
want to retain incremental (modular) type checking}. Notice that, the
modularity here in Coq is a bit different from other languages, because
Coq supports \textit{interactive} theorem proving, so we actually need
\textit{statement-wise} incremental type checking, not only for avoiding
re\-compilation, but also to enable immediate feedback and incremental
type-checking for the Coq developer. We don't want our family to ruin
this conventional routine of interactive development.
\textlabel{Req~(0B)}{langdesign-req0b}~\textbf{we also want to keep the
computational ability of Coq when using families}. Coq, based on constructive
logic, can consider proof as programs. 
We don't want our Family facility to ruin this when incorporated with 
the remaining parts of Coq: 
the developers should be able to project fields of families 
as first-class value just like how they can project fields of Coq's Module.

% We want to reason about fields in a family.
\textlabel{Req (1)}{langdesign-req1}:\textbf{ we want to be able to
reason about fields in a family}, just like in a module, where a field
can reason about its former fields. More generally, we want to allow a
later field to be type-dependent on the former fields. In the example of
\cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
this feature.

% we want extensible inductive type
\textlabel{Req (2)}{langdesign-req2}:\textbf{ we need extensible
inductive types}, to extend "term" and "ty" in \cref{fig:STLC-example}.
Extensible inductive types are not supported by \citet{zm2017}, so
we need further consideration about it on both implementation and
metatheory.

\textlabel{Req (3)}{langdesign-req3}: \textbf{we also want extensible
pattern matching and induction reasoning}, to extend "subst" in
\cref{fig:STLC-example}.
There are actually two kinds of ``pattern matching'', one for data and
the other for induction reasoning, i.e., one uses the eliminator to
\mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
extensible pattern matching can be easily expressed with family
extension---we just aggregate all case handlers of pattern matching
into one family (as a bunch of functions), and then one family can
encode one pattern matching, and family inheritance can express adding
case handlers. Then we just need to introduce a primitive that will
``wrap'' that family into a recursive function. Induction reasoning can
be handled in an almost identical way.

% we want to be able to reason pattern matching as well
However, there is still difference between data recursion and induction
reasoning, because the former one is \textit{proof-relevant}. This
difference leads to another issue: \textlabel{Req (4)}{langdesign-req4}:
\textbf{we need to reason about (the computation about) extensible
pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
When proving "subst_lemma", we have to know "subst" is invariant on
"tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
information requires exposing \textit{computational rules} from the
recursors.\YZ{Seems like what's missing here is to
convince the reader why exposing the definition of subst could be a
problem for family polymorphism.}\EDJreply{Actually I don't see why I need to emphasize this "as a problem". This is just a requirement like any other requirements earlier. I don't know how to show its difficulty at this early stage.}\YZreply{There is apparently a tension between the need to expose definitional equality and the ability to override. Looking at the bigger picture, I think this section reads better if phrased as discussing challenges posed by adding family polymorphism to a proof assistant. I have some initial ideas that I need to discuss with you.}

% we want tactic programming and certain level of automation
Finally, \textlabel{Req (5)}{langdesign-req5}: \textbf{we desire tactic
programming} to relieve us from the hassle of direct manipulation of proof terms.

>>>>

\subsection{Challenges}

We want our metatheory mechanization to allow adding new function/proving new lemmas, and adding new constructors for inductive type at the same time  (Expression Problem~\cite{wadler-ep}). Coq support the former but not the later---because when an existent inductive type is extended with new constructor, we have to modify the pattern matching in the existent functions. This motivates us to add inheritance/Family Polymorphism into Coq.

\begin{figure}[!htb]
  \begin{minipage}[t]{0.47\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=||]{Coq}
Family STLC.|\YZ{Does minted support adding extra keywords such as 'Family' and 'extends'?}|
Inductive ty : Set :=
  | unit : ty
  | arrow : ty -> ty -> ty.
Inductive term : Set := 
  | var : id -> term 
  | lam : id -> term -> term ...
  | tunit : term 
Fixpoint subst 
  : id -> term -> term := ...
Inductive has_type 
  : context -> term -> type := ...
Proposition subst_lemma :
  has_type (Γ; x ↦ A) t T ->
  has_type Γ u A ->
  has_type Γ (subst x u t) T.
Inductive step : term -> term -> Prop 
  := ...
(* ... and more, end with type safety *)
EndFamily.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.47\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family STLC_bool extends STLC.
Extend Inductive ty : Set :=
  | bool : ty.

Extend Inductive term : Set := 
  | tt : term | ff : term 
  | tif : term -> term -> term -> term

Extend Fixpoint subst (* Need to handle new term *)
  : id -> term -> term := ...
Extend Inductive has_type (* .. and new ty too *)
  : context -> term -> type := ...
Extend Proposition subst_lemma :
  ... (* Need to prove extra cases *)

@@

However, to add inheritance into a dependent type system like Coq is not simple and there are various challenges and tension we need to resolve.

\textlabel{Challenge (1)}{chg:extensible-inductive-type}~\textbf{Extensible Inductive Type vs. Exhaustiveness of Inductive Reasoning}. Unlike most general purposed programming languages, Coq, as a dependent type theory, is a total language and thus each pattern matching cases must fully ``cover'' an inductive type. This process is \textit{dynamic} for inheritance---when inhertiance extend the inductive type in the children family but the user forget to extend the corresponding inductive reasoning, our compiler need to signal an error correspondingly, or better---guide the user to fill the incomplete inductive reasoning. For example, in \cref{fig:STLC-example} when "term" is extended with "tt" in the children family "STLC_bool", our compiler need to make sure "subst" and "subst_lemma" are also extended correctly otherwise signaling an error.
Additionally, our compiler must support direct inheritance (and overriding) of the old case handlers, to waive the boilerplate code.

\textlabel{Challenge (2)}{chg:definition-relevant-reasoning}~\textbf{Definition-awared Reasoning vs. Overridability}. In family polymorphism for general purposed programming language, overriding is allowed to happen anywhere. But the problem becomes much more subtle in the case of dependent type---the later fields can have a type dependent on the former fields! Consider a small example where we have consecutive fields \mintinline{Coq}{{.. a : nat := 1; a-def : Eq a 1 := eq_refl; ..}}, should we support override "a"? If we do support it, how can our compiler detect that "a-def" is not inheritable and signal an error when such thing happens?

A similar example is about "subst" and "subst_lemma" in \cref{fig:STLC-example}. For "subst_lemma" to hold throughout two families, we shouldnot allow overriding on the definition of "subst". For example, we should still stick with the fact that "subst tunit x y = tunit", as this fact is used when proving "subst_lemma" and we want this lemma to be inherited to the children family.

These two examples both show the incompatibility between exposing the concrete definition of a field for reasoning (this second example even involves the definition of pattern matching) and overridability. We need to be careful about this tension when designing the language.

\textlabel{Challenge (3)}{chg:consistency}~\textbf{Consistency vs. Self Parametrization}. The metatheoretical formulation in~\citet{zm2017} for family polymorphism is very natural as they consider extension of the family as the concatenation of a new linkage. However, this formulation is not directly adaptable for dependent type theory because it can easily encode non-termination (mutual calling of two function member) sabotaging consistency by the presence of "self". This conflict between self parametrization and consistency will be another challenge---we need change the original formulation of family and inheritance for dependent type theory.

\textlabel{Challenge (4)}{chg:software-engineering}~\textbf{Engineering and Ecosystem}.  Not only for the metatheory, we also want our implementation of family to have good compatibility with the rest of Coq. For example, we cannot compromise \textbf{tactic programming} as it is already a widely accepted part of Coq programming. Additionally, the interactive development style requires \textbf{incremental type checking}---our compiler has to be able to type check each Coq Vernacular command. What's more, we want our family ``term'' to incorporate Coq \textbf{without breaking its other functionality}. For example, the user may want to project the family member anywhere inside an arbitrary Gallina term, that is later being computed and normalized. However, families themselves can be second class citizens just like \citet{zm2017}.

\YZ{I was thinking if the paper should start by talking about the functional encoding of inheritance (namely exposing a self parameter to allow late binding, and later taking the fixpoint of the extensible module when it needs to be accessed outside the family).  I think most readers will benefit from a recap of this encoding.\\
Coq is total but that doesn’t prevent Coq from having a Fixpoint keyword (plus you proved your language is consistent).
}

% \subsection{Design Space}

% When adapting Family Polymorphism into dependent types, we choose to
% focus only on the essence of family (and inheritance)
% structure in \citet{zm2017}, and thus a lot of unrelated features, like
% Interface, will be removed. In this case, it will look like module with
% late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
% need it to have good compatibility with inductive types, because we
% don't want to retain the idiomatic Coq programming experience
% %\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness. I rephrase it now, please check.}
% (and its tactic programming).

% Let's start with a basic example---STLC and its extension in
% \cref{fig:STLC-example}---to consider what kind of features are required
% and how much of them can be supported by family polymorphism.
% add the example code here

% First and foremost, we don't want to throw away some basic good
% properties of Coq. \textlabel{Req~(0A)}{chg:software-engineering}~\textbf{we
% want to retain incremental (modular) type checking}. Notice that, the
% modularity here in Coq is a bit different from other languages, because
% Coq supports \textit{interactive} theorem proving, so we actually need
% \textit{statement-wise} incremental type checking, not only for avoiding
% re\-compilation, but also to enable immediate feedback and incremental
% type-checking for the Coq developer. We don't want our family to ruin
% this conventional routine of interactive development.
% \textlabel{Req~(0B)}{chg:software-engineeringb}~\textbf{we also want to keep the
% computational ability of Coq when using families}. Coq, based on constructive
% logic, can consider proof as programs. 
% We don't want our Family facility to ruin this when incorporated with 
% the remaining parts of Coq: 
% the developers should be able to project fields of families 
% as first-class value just like how they can project fields of Coq's Module.

% We want to reason about fields in a family.
% \textlabel{Req (1)}{chg:definition-relevant-reasoning}:\textbf{ we want to be able to
% reason about fields in a family}, just like in a module, where a field
% can reason about its former fields. More generally, we want to allow a
% later field to be type-dependent on the former fields. In the example of
% \cref{fig:STLC-example}, "subst", "has_type", "subst_lemma" all require
% this feature.

% we want extensible inductive type
% \textlabel{Req (2)}{chg:extensible-inductive-type}:\textbf{ we need extensible
% inductive types}, 
% to extend "term" and "ty" in \cref{fig:STLC-example}.
% Extensible inductive types are not supported by \citet{zm2017}, so
% we need further consideration about it on both implementation and
% metatheory.

% \textlabel{Req (3)}{chg:extensible-inductive-type}: \textbf{we also want extensible
% pattern matching and induction reasoning}, to extend "subst" in
% \cref{fig:STLC-example}.
% There are actually two kinds of ``pattern matching'', one for data and
% the other for induction reasoning, i.e., one uses the eliminator to
% \mintinline{Coq}{Set} or \mintinline{Coq}{Type}, and the other uses the
% eliminator to \mintinline{Coq}{Prop}. Luckily, in this setting,
% extensible pattern matching can be easily expressed with family
% extension---we just aggregate all case handlers of pattern matching
% into one family (as a bunch of functions), and then one family can
% encode one pattern matching, and family inheritance can express adding
% case handlers. Then we just need to introduce a primitive that will
% ``wrap'' that family into a recursive function. Induction reasoning can
% be handled in an almost identical way.

% we want to be able to reason pattern matching as well
% However, there is still difference between data recursion and induction
% reasoning, because the former one is \textit{proof-relevant}. This
% difference leads to another issue: \textlabel{Req (4)}{chg:definition-relevant-reasoning}:
% \textbf{we need to reason about (the computation about) extensible
% pattern matching}, just like "subst_lemma" in \cref{fig:STLC-example}.
% When proving "subst_lemma", we have to know "subst" is invariant on
% "tt", i.e., \mintinline{Coq}{(subst i x tt) = tt}. This kind of
% information requires exposing \textit{computational rules} from the
% recursors.
\YZ{Seems like what's missing here is to
convince the reader why exposing the definition of subst could be a
problem for family polymorphism.}\EDJreply{Actually I don't see why I need to emphasize this "as a problem". This is just a requirement like any other requirements earlier. I don't know how to show its difficulty at this early stage.}\YZreply{There is apparently a tension between the need to expose definitional equality and the ability to override. Looking at the bigger picture, I think this section reads better if phrased as discussing challenges posed by adding family polymorphism to a proof assistant. I have some initial ideas that I need to discuss with you.}

% we want tactic programming and certain level of automation
% Finally, \textlabel{Req (5)}{chg:software-engineering}: \textbf{we desire tactic
% programming} to relieve us from the hassle of direct manipulation of proof terms.
